---
title: "Exercise solutions: Section 2.8"
author: "Rob J Hyndman & George Athanasopoulos"
output: html_document
---
```{r, echo = F, message = F}
require(fpp)
```

## Exercise 2.1

### Monthly total of people on unemployed benefits in Australia (January 1956–July 1992)

```{r}
require(fpp) # fpp package needs to be loaded once in the beginning of every R session
lambda = BoxCox.lambda(dole)
plot(BoxCox(dole, lambda), xlab = "Time", 
  ylab = paste("BoxCox(# people,", round(lambda, 2), ")"))
```

The data was transformed using Box-Cox transformation with parameter $\lambda=`r round(lambda,2)`$.

### Monthly total of accidental deaths in the United States (January 1973–December 1978)

```{r}
plot(usdeaths, xlab = "Time", ylab = "# deaths")
```

### Quarterly production of bricks (in millions of units) at Portland, Australia (March 1956–September 1994)

```{r}
lambda = BoxCox.lambda(bricksq)
plot(BoxCox(bricksq, lambda), xlab = "Time", 
     ylab = paste("BoxCox(# mln bricks,", round(lambda, 2), ")"))
```

The data was transformed using Box-Cox transformation with parameter $\lambda=`r round(lambda,2)`$.


\newpage


## Exercise 2.2

Time plot of the Dow Jones index:

```{r}
plot(dowjones, xlab = "Time", ylab = "Value $")
```

Forecasts using the drift method:

```{r}
plot(rwf(dowjones, drift = TRUE, h = 20), xlab = "Time", ylab = "Value $", main = "")
```

Graphed forecasts are identical to extending the line drawn between the first and last observations:

```{r}
plot(rwf(dowjones, drift = TRUE, h = 20), xlab = "Time", ylab = "Value $", main = "")
slope = (tail(dowjones, 1) - head(dowjones, 1)) / (length(dowjones) - 1)
intercept = head(dowjones, 1) - slope # Since time starts from 1
abline(intercept, slope, lty = 2, col = "red")
```

Some other benchmark methods:

```{r}
plot(rwf(dowjones, drift=TRUE, h=20, level=0), xlab="Time", ylab="Value $", main="")
lines(naive(dowjones, h=20, level=0)$mean, xlab="", ylab="", main="", col="green")
lines(meanf(dowjones, h=20, level=0)$mean, xlab="", ylab="", main="", col="red")
# "$mean" after the function calls above is used
# to extract vector of the forecasted values from the function output
legend("topleft",
  legend = c("Random walk with drift", "Random walk without drift", "Mean forecast"),
  col = c("blue", "green", "red"), lty=1)
```

Random Walk Forecast (`rwf`) method with and without drift might be the best forecasting methods.


\newpage


## Exercise 2.3

Plot of daily closing IBM stock prices (data set ibmclose):

```{r}
plot(ibmclose, xlab = "Time", ylab = "Value $", main = "")
```

Splitting the data into a training set of 300 observations and a test set of 69 observations:

```{r}
trainingSet = window(ibmclose, 1, 300)
testSet = window(ibmclose, 301, 369)
```

Benchmark methods are used to forecast the training set. The results are compared on the test set.

```{r}
rwfForecast = rwf(trainingSet, h = 69)$mean
rwfWithDriftForecast = rwf(trainingSet, drift = TRUE, h = 69)$mean
meanfForecast = meanf(trainingSet, h = 69)$mean
```
```{r, comment = ""}
# Comparing the above forecasts
accuracy(rwfForecast, testSet)
accuracy(rwfWithDriftForecast, testSet)
accuracy(meanfForecast, testSet)
```

According to RMSE (Root Mean Square Error) the best forecast is provided by the Random Walk Forecast with Drift method.  
If MAE (Mean Absolute Error) or MAPE (Mean Absolute Percentage Error) are considered, they give the same result.


\newpage


## Exercise 2.4

Plot of monthly sales of new one-family houses in the USA, Jan 1973 – Nov 1995 (data set `hsales`):

```{r}
plot(hsales, xlab = "Time", ylab = "Sales", main = "")
```

Splitting the hsales data set into a training set and a test set, where the test set is the last two years of data.

```{r}
trainingSet = window(hsales, end = c(1993,12))
testSet = window(hsales, start = c(1994,1))
```

Benchmark methods are used to forecast the training set. The results are compared on the test set.

```{r}
rwfForecast = rwf(trainingSet, h = 23)$mean
rwfWithDriftForecast = rwf(trainingSet, drift = TRUE, h = 23)$mean
meanfForecast = meanf(trainingSet, h = 23)$mean
snaiveForecast = snaive(trainingSet, h = 23)$mean
```
```{r, comment = ""}
# Comparing the above forecasts
accuracy(rwfForecast, testSet)
accuracy(rwfWithDriftForecast, testSet)
accuracy(meanfForecast, testSet)
accuracy(snaiveForecast, testSet)
```

According to RMSE (Root Mean Square Error) the best forecast is provided by the Seasonal Naive method.  
If MAE (Mean Absolute Error) or MAPE (Mean Absolute Percentage Error) are considered, they give the same result.

